
# Getting and Cleaning Data course Project

## 1. Downloading and unzip the Data 

```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./data/Dataset.zip")
unzip(zipfile="./data/Dataset.zip",exdir="./data")
```

Loading some R packages 

```{r,recursive =TRUE}
library(dtplyr)
library(tidyr)
```

## 2. Getting the List of Files

```{r}
filepath<- file.path("./data" , "UCI HAR Dataset")
list<-list.files(filepath, recursive=TRUE)
list
```


### Reaading activity Variables from Data files

```{r}
dataActivityTest  <- read.table(file.path(filepath, "test" , "Y_test.txt"),header = FALSE)
dataActivityTrain <- read.table(file.path(filepath, "train", "Y_train.txt"),header=FALSE)
```

### Reading Subject files 

```{r}
dataSubjectTrain <- read.table(file.path(filepath, "train", "subject_train.txt"),header=FALSE)
dataSubjectTest  <- read.table(file.path(filepath, "test" , "subject_test.txt" ),header=FALSE)
```

###Reading Features files
```{r}
dataFeaturesTest  <- read.table(file.path(filepath, "test" , "X_test.txt" ),header = FALSE)
dataFeaturesTrain <- read.table(file.path(filepath, "train", "X_train.txt"),header = FALSE)
```


```{r}
str(dataActivityTest)
```

## 3. Merging the training and the testing  datasets to Create one dataset

```{r}
dataSubject <- rbind(dataSubjectTrain, dataSubjectTest)
dataActivity<- rbind(dataActivityTrain, dataActivityTest)
dataFeatures<- rbind(dataFeaturesTrain, dataFeaturesTest)
```

### b.giving names to the variables

```{r}
names(dataSubject)<-c("subject")
names(dataActivity)<- c("activity")
dataFeaturesNames <- read.table(file.path(filepath, "features.txt"),head=FALSE)
names(dataFeatures)<- dataFeaturesNames$V2
```


### c.Merging Data sets
```{r}
dataCombine <- cbind(dataSubject, dataActivity)
Data <- cbind(dataFeatures, dataCombine)
head(Data)
```
##4. Extracting only the measurements on the mean and standard deviation for each measurement

```{r}
subdataFeaturesNames<-dataFeaturesNames$V2[grep("mean\\(\\)|std\\(\\)", dataFeaturesNames$V2)]
selectedNames<-c(as.character(subdataFeaturesNames), "subject", "activity" )
Data<-subset(Data,select=selectedNames)
head(Data)
```
### Naming the actvites in the dataset

```{r}
activityLabels <- read.table(file.path(filepath, "activity_labels.txt"),header = FALSE)
Data$activity<- factor(Data$activity,labels=as.character(activityLabels$V2))
head(Data$activity,30)
```


```{r}
names(Data)<-gsub("^t", "time", names(Data))
names(Data)<-gsub("^f", "frequency", names(Data))
names(Data)<-gsub("Acc", "Accelerometer", names(Data))
names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
names(Data)<-gsub("Mag", "Magnitude", names(Data))
names(Data)<-gsub("BodyBody", "Body", names(Data))
```

##5. Creating a second independent tidy Data set
```{r}
library(dplyr);
finaldata<-aggregate(. ~subject + activity, Data, mean)
finaldata<-finaldata[order(finaldata$subject,finaldata$activity),]
write.table(finaldata, file = "mytidydata.txt",row.name=FALSE)
```



