

# Getting and Cleaning Data course Project

## 1. Downloading and unzip the Data 

```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./data/Dataset.zip")
unzip(zipfile="./data/Dataset.zip",exdir="./data")
```

```{r}
library(dplyr)
```

## 2. Getting the List of Files

```{r}
filepath<- file.path("./data" , "UCI HAR Dataset")
list<-list.files(filepath, recursive=TRUE)
list
```


## Creating y Data set 
```{r}
y_test  <- read.table(file.path(filepath, "test" , "Y_test.txt"),header = FALSE)
y_train <- read.table(file.path(filepath, "train", "Y_train.txt"),header=FALSE)
y_dataset<-rbind(y_train, y_test)
str(y_dataset)
```


## Creating subject data set
```{r}
sub_train <- read.table(file.path(filepath, "train", "subject_train.txt"),header=FALSE)
sub_test  <- read.table(file.path(filepath, "test" , "subject_test.txt" ),header=FALSE)
sub_dataset<-rbind(sub_train,sub_test)
str(sub_dataset)
```

## Creating x data set
```{r}
x_test  <- read.table(file.path(filepath, "test" , "X_test.txt" ),header = FALSE)
x_train <- read.table(file.path(filepath, "train", "X_train.txt"),header = FALSE)
x_dataset<-rbind(x_train, x_test)
```

# Extracting only the measurements on the mean and standard deviation for each measurement 

```{r}
features<-read.table(file.path(filepath, "features.txt"),head=FALSE)
mean_std <- grep("-(mean|std)\\(\\)", features[, 2])
mn_std<-features$V2[grep("mean\\(\\)|std\\(\\)", features$V2)]
head(mean_std)
head(mn_std)
```

## Subset and correcting the desired columns 

```{r}
x_data <- x_dataset[, mean_std]
names(x_data) <- features[mean_std, 2]
names(x_data)
```


### Use descriptive activity names to name the activities in the data set


```{r}
activitylabels <- read.table(file.path(filepath, "activity_labels.txt"),header = FALSE)
y_dataset[, 1] <- activitylabels[y_dataset[, 1], 2]
names(y_dataset) <- "activity"
```

## Appropriately label the data set with descriptive variable names

```{r}
names(sub_dataset) <- "subject"
```

### Creating a single data set 

```{r}
dataset <- cbind(x_data, y_dataset, sub_dataset)
```

## Create a second, independent tidy data set with the average of each variable for each activity and each subject

```{r}
newdata <- ddply(dataset, .(subject, activity), function(x) colMeans(x[, 1:65]))

write.table(newdata, "mynewdata.txt", row.name=FALSE)
```




